---
title: "Sergio Mora ST 558 Project 1"
author: "Sergio Mora"
date: "10/2/2021"
output:
  github_document:
    toc: true
    toc_depth: 1
---

# Required Packages for this Vingette

  - The `lubridate` for date manipulation
  - The `jsonlite` to pull date from API's
  - The `tidyverse` to utilize infix functions such as `%>%`
  - The `countrycode` package to join the country of origin to the continent it belongs in
  - The `knitr` package to show our data in a way that is more appialing
  - The `reshape2` package to reshape my data. Although this can be achieved with pivot_longer/pivot_wider `reshape2` is easier to use
```{r, message = FALSE, warning=FALSE}
library(lubridate)
library(jsonlite)
library(tidyverse)
library(countrycode)
library(knitr)
library(reshape2)
```

# Creating New Functions / Data Manipulation
We are creating two functions:

  - `referrence_table_lookup()`: which lets us pull any countries covid data based on our input. This will be helpful if we want to loop through these countries to get similar analysis across the board.
  
  - `referrence_table_clean()`: which pulls in `referrence_table_lookup()` and cleans it up to how we want to see the data for our plots down below. We also create a few new variables which we will use for out Exploratory Data Analysis (EDA).

```{r, message=FALSE}
referrence_table <- fromJSON(
    "https://api.covid19api.com/countries"
    )


# This gives us the continent that each country belong in. There are a few exception that this function does not count for, we will not acocunt for these manually since we are not looking at these individually and because my geography skills are awful.
referrence_table <- referrence_table %>% mutate(continent = countrycode(sourcevar = referrence_table[, "Slug"], origin = "country.name", destination = "continent"))

#Creating a function to pull in any country data.
referrence_table_lookup <- function(x,type){
country_clean <-   switch (type,
    Country = referrence_table %>% filter(Country == x) %>% select(Slug),
    Slug = referrence_table %>% filter(Slug == x) %>% select(Slug),
    ISO2 = referrence_table %>% filter(ISO2 == x) %>% select(Slug)
  )
return(fromJSON(paste0("https://api.covid19api.com/live/country/",country_clean,"/status/confirmed"))
)
}


referrence_table_clean <- function(x,type){
referrence_table_lookup(x,type) %>% filter(Date > as.Date("2021-01-01")) %>% group_by(Country,Date) %>% summarise(sum_of_Confirmed = sum(Confirmed),sum_of_Deaths = sum(Deaths),sum_of_Active = sum(Active))  %>% mutate(New_Confirmed = (sum_of_Confirmed - lag(sum_of_Confirmed)),New_Deaths = (sum_of_Deaths - lag(sum_of_Deaths)),New_Active = (sum_of_Active - lag(sum_of_Active)),Date = ymd_hms(Date))
  }
```

# Contingency Table
We want to see how many individual countries we have per continent. Due to vast difference in population this doesn't tell us a huge amount with regards to any preditive models we could build but it's nice to know. There are a few terretories that are not accounted.

```{r}
table(referrence_table$continent) %>% kable(col.names = c("Continent", "Frequency"))
```


# Creating New Data Sets
Now that we have created two functions that will allow us to pull data directly from the API for any country available we will now create a few datasets to review. I live in the U.S. but I was born and raised in Venezuela. For this reason I am currous how Venezuela compares to the U.S. in covid cases. For good measure we will also through in France and Colombia, these are countries that I have spent some time in growing up. There is no need to further manipulate our data since we took care of that in our functions above.
```{r, message=FALSE}
Full_data <- bind_rows(referrence_table_clean("US","ISO2"),referrence_table_clean("venezuela","Slug"),referrence_table_clean("France","Country"),referrence_table_clean("Colombia","Country"))
```


# Numerical Summaries / Graphs
```{r}
Full_data %>% summarise(avg_new_confirmed = round(mean(New_Confirmed, na.rm = TRUE)),avg_new_Deaths= round(mean(New_Deaths, na.rm = TRUE)))

g <- ggplot(Full_data, aes(x = Date))
g + geom_line(aes(y = New_Confirmed)) + labs(title = "New Confimed cases by country", ylab = "Newly Confirmed Cases") + facet_wrap(~ Country)


#Further manipulating the data


g + geom_line(aes(y = New_Deaths)) + facet_wrap(~ Country) + labs(title = "New Confimed cases by country", ylab = "Newly Confirmed Deaths")

```

# Plots

```{r, warning=FALSE, message=FALSE}
g <- ggplot(Full_data)
g + geom_point(aes(x = New_Confirmed, y = New_Deaths, color = Country))
g + geom_boxplot(aes(x = Country, y = New_Deaths, color = Country)) + theme(axis.text.x = element_text(angle = -45))


Summary_data <- Full_data %>% group_by(Country) %>% summarise(Avg_Deaths = mean(New_Deaths, na.rm = TRUE), Avg_Confirmed = mean(New_Confirmed, na.rm = TRUE))
q <- ggplot(melt(Summary_data), aes(x = Country, y = value, fill = variable, color = variable))
q + geom_bar(stat = "identity", position = "dodge") + ylab("Average cases Confirmed or Dead") + theme(axis.text.x = element_text(angle = -45))

```



```{r}
q <- ggplot(Summary_data, aes(x = Country))
q + geom_bar(aes(y = Avg_Confirmed), stat = "identity") + geom_bar(aes(y = Avg_Deaths), stat = "identity")






q + geom_bar(aes(y = Avg_Confirmed), stat = "identity") + geom_bar(aes(y = Avg_Deaths), stat = "identity")

```

